Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots

### 2024-11-12 paper/test.Rmd
interactive maps revised

# Create a new column 'magnitude' that counts the total number of occurrences for each value in 'hood_num'
map_data <- map_data %>% 
  group_by(hood_num) %>%       # Groups the data by 'hood_num'.
  mutate(magnitude = n()) %>%  # Adds a 'magnitude' column, which counts the number of rows in each group.
  ungroup()                    # Ungroups the data, making it ready for further manipulations.

# Modify the data frame, categorizing magnitude column
map_data <- map_data %>% 
  mutate(
    mag.level = cut(magnitude, c(100, 500, 1000, 1500),
                    labels = c('> 100 & <=500', '>500 & <=1000', '>1000 & <=1500'))
    # Creates a 'mag.level' column that categorizes 'magnitude' into three levels based on specified ranges.
  ) %>%
  split(.$mag.level)  # Splits the dataset into a list of data frames by 'mag.level'.

@@

# Add the markers to the map for each magnitude level
names(map_data) %>%
  walk(function(df) {
    l <<- l %>% 
      addMarkers(
        data = map_data[[df]], lng = ~longitude, lat = ~latitude,
        label = ~as.character(magnitude), popup = ~as.character(magnitude),
        group = df,
        clusterOptions = markerClusterOptions()
      )
  })
# Iterates over the names of the 'map_data' list (which are the different 'mag.level' categories).
# For each category, it adds markers to the map:
# - `data = map_data[[df]]`: Uses the split data for the current magnitude level.
# - `lng = ~longitude, lat = ~latitude`: Sets longitude and latitude for marker placement.
# - `label` and `popup`: Display the 'magnitude' value when markers are clicked or hovered over.
# - `group = df`: Groups the markers by 'mag.level'.
# - `clusterOptions = markerClusterOptions()`: Enables clustering of markers for better visualization.

@@

# Load your dataset (adjust the file path accordingly)
map_data <- read_csv(here("data", "map_by_neighbourhoods_data.csv"), show_col_types = FALSE)

# Create a new column 'magnitude' that counts the total number of occurrences for each value in 'hood_num'
map_data <- map_data %>% 
  group_by(hood_num) %>% 
  summarise(
    magnitude = n(),
    longitude = mean(longitude),
    latitude = mean(latitude)
  ) %>% 
  ungroup()

@@

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "highway", value = "primary") %>%
  add_osm_feature(key = "highway", value = "secondary") %>%
  add_osm_feature(key = "highway", value = "tertiary") %>%
  add_osm_feature(key = "highway", value = "residential") %>%
  osmdata_sf()

@@

# Create ggplot object with the dataset and Toronto map as a background
ggplot_map <- ggplot() +
  geom_sf(data = streets, color = "gray", size = 0.3, alpha = 0.6) +  # Add OSM street map layer
  geom_point(data = shared_data, aes(x = longitude, y = latitude, color = magnitude, size = magnitude)) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "Toronto Neighbourhood Data", x = "Longitude", y = "Latitude") +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Convert ggplot object to interactive plotly object
plotly_map <- ggplotly(ggplot_map) %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    selected = attrs_selected(line = list(color = "black"))
  )

>>>>

# Modify the data frame, categorizing case_num column
map_data <- map_data %>% 
  mutate(
    mag.level = cut(case_num, c(100, 500, 1000, 1500),
                    labels = c('> 100 & <=500', '>500 & <=1000', '>1000 & <=1500'))
    # Creates a 'mag.level' column that categorizes 'case_num' into three levels based on specified ranges.
  ) %>%
  split(.$mag.level)  # Splits the dataset into a list of data frames by 'mag.level'.

@@

# Add the markers to the map for each case_num level
names(map_data) %>%
  walk(function(df) {
    l <<- l %>% 
      addMarkers(
        data = map_data[[df]], lng = ~longitude, lat = ~latitude,
        label = ~paste0("Hood: ", hood_name, " | Stolen Place: ", stolen_place, " | Cost: ", cost, " | Status: ", status),
        popup = ~paste0(
          "Hood: ", hood_name, "<br>",
          "Stolen Place: ", stolen_place, "<br>",
          "Cost: ", cost, "<br>",
          "Status: ", status, "<br>",
          "Case in area: ", case_num
        ),
        group = df,
        clusterOptions = markerClusterOptions()
      )
  })
# Iterates over the names of the 'map_data' list (which are the different 'mag.level' categories).
# For each category, it adds markers to the map:
# - `data = map_data[[df]]`: Uses the split data for the current case_num level.
# - `lng = ~longitude, lat = ~latitude`: Sets longitude and latitude for marker placement.
# - `label`: Displays the 'case_num' value when markers are hovered over.
# - `popup`: Displays additional information (hood_num, stolen_place, cost, status, and case_num) when a marker is clicked.
# - `group = df`: Groups the markers by 'mag.level'.
# - `clusterOptions = markerClusterOptions()`: Enables clustering of markers for better visualization.

@@

# Load your dataset (adjust the file path accordingly)
map_data <- read_csv(here("data", "map_by_area_data.csv"), show_col_types = FALSE)

@@

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "boundary", value = "administrative") %>%
#  add_osm_feature(key = "highway", value = "primary") %>%
#  add_osm_feature(key = "highway", value = "secondary") %>%
#  add_osm_feature(key = "highway", value = "tertiary") %>%
#  add_osm_feature(key = "highway", value = "residential") %>%
  osmdata_sf()

@@

# Create ggplot object with the dataset and Toronto map as a background
ggplot_map <- ggplot() +
  geom_sf(data = streets, color = "lightgray", size = 0.3, alpha = 0.6) +  # Add OSM street map layer
  geom_point(data = shared_data, aes(x = longitude, y = latitude, color = case_num, size = case_num, 
                                     text = paste("Total Cases: ", case_num, " | Area Name: ", area_name))) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "Toronto Neighbourhood Data by Area", x = "Longitude", y = "Latitude", color = "Number of Cases", size = "Total Cases") +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_continuous(labels = scales::comma)

# Convert ggplot object to interactive plotly object
plotly_map <- ggplotly(ggplot_map, tooltip = "text") %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    selected = attrs_selected(line = list(color = "black"))
  )
